Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2018-06-17 2018/Journal of Supercomputing/JoS_Paper.tex
Similarity correction

\Russian
Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными.

Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the criteria is indeterminate.

@@

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are given in \cite{Sergeyev2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the domains 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

@@

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
\end{itemize}
The pair of values $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ is a \textit{trial result}.

@@

\Russian
В статье \cite{Barkalov2017} приведено сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы. Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции. Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области). Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.

@@

The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function's divided differences. However, a partial loss of information on the nearness of the points in the multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions. As a result, when using the mappings like a Peano curve, the images $y', y''$, which are close to each other in the $N$-dimensional space can correspond to the preimages $x', x''$, which can be far away from each other in the interval $[0,1]$. This property results in excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

Another scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}. The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme. In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations might appear to be redundant. In this case, we can select only a part of all rotations. As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes. Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$. 
\Russian
Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов

With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by solving in parallel the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor/node. The trial results at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the results of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.

@@

	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k)\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The trial terminates at the point as soon as the first violation of any constraint occurs. In the case when $y^k$ is a feasible one, the trial includes the computation of the values of all problem functions. In this situation, the index is set to $\nu=m+1$. The triplet
\[
y^s(x^k), \; \nu=\nu(x^k), \; z^k=g_\nu(y^s(x^k))
\]
is the result of the trial at the point $x^k$.

\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same results
\[
 \nu(x^{k_1}) = ... = \nu(x^{k_S}) = \nu(x^k),
\]
\[
 g_\nu(y^1(x^{k_1})) = ... = g_\nu(y^S(x^{k_S})) = z^k.
\]

\item Inform the rest of processors about the trial results at the point $y^k$.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. 
\Russian
Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области. При этом можно варьировать объем допустимой области, а также число ограничений, активных в точке минимума.

@@

The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$. The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain. For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).

Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod. The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The CPU had 8 cores, i.e. each node had a total of 16 cores.

The average time and the number of iterations, which were required to solve the problems of the series at various parallelization parameters are given in Tables 1 and 2. Here, \textit{Node(core)} are the numbers of employed  nodes and cores per node, respectively.

@@

The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes). In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.

@@

The parallel index method for solving constrained global optimization problems considered in this paper offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed. This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer. 
\Russian
Что-то про применение алгоритма. 
%The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.

@@

\bibitem{Balandin2017}
Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)

>>>>

\Russian
Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными. Например, моделируемым объектом является некоторое электронное устройство, и минимизируется время его срабатывания. Однако время срабатывания является неопределенным, если при данном сочетании параметров устройство вообще не работает.

%Another example is the optimal control problem described through system of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the criteria is indeterminate.

@@

A constrained global optimization problem can be formulated as follows
%Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (hereafter denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy the Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priori unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is assumed that the functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval $[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are considered in \cite{Sergeyev2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined and computable in the subranges 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

@@

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping $y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from (\ref{nu}). 
\end{itemize}
The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial outcome}.

@@

\Russian
Сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы - приведено в статье \cite{Barkalov2017}. Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции. Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области). Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.

@@

The reduction of the multidimensional problems to the one-dimensional ones using evolvents has such important properties as the continuity and preservation of boundedness of function's divided differences. However, a partial loss of information on the nearness of the points in the multidimensional space occurs in this case since a point $x \in [0,1]$ has only the left and the right neighbors while the corresponding point $y(x) \in R^N$ has the neighbors in $2N$ directions. As a result, when using the mappings like a Peano curve, the close images $y', y'' \in D$ can correspond to the remote preimages $x', x'' \in [0,1]$. This property results in excess computations since several limit points $x', x''$ of the trial sequence generated by the index method in the interval $[0,1]$ can correspond to a single limit point $y$ in the $N$-dimensional space.

One of the possible ways to overcome this disadvantage consists in using the multiple mapping $Y^S(x)=\{ y^1 (x),...,y^S(x)\}$ instead of the single evolvent $y(x)$. To construct the set $Y^S(x)$, different approaches can be used. 
%For example, in \cite{Strongin2000} a scheme was implemented, according to which each evolvent $y^s(x)$ from $Y^S(x)$ is constructed as a result of shifting the original evolvent $y(x)$ along the main diagonal of the hypercube $D$. The set of Peano curves thus constructed allows one to obtain close preimages $x', x''$ from the interval $[0,1]$ for the evolvent $y^s(x), 1\leq s \leq S$, for any close multidimensional images $y', y''$ from $D$, which differ only in one coordinate.

%Проверить
For example, a scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}. The building of a set of Peano curves by rotation of the evolvents around the coordinate origin is a distinctive feature of this scheme. In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations might appear to be redundant. In this case, we can select only a part of all rotations. As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes. Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.
\Russian
Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов.

With the use of multiple mapping, it is possible to solve the initial problem (\ref{problem}) by solving in parallel the problems
\[
\min \left\{\varphi(y^s(x)): x \in [0,1], \; g_i(y^s(x))\leq 0, \; 1 \leq i \leq m\right\},\; 1\leq s \leq S.
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor/node. The trial outcome at the point $x^k$ obtained for the problem being solved by particular processor are interpreted as the outcomes of the trials in the remaining problems (at the corresponding points $x^{k_1},...,x^{k_S}$). In this approach, a trial at the point $x^k \in [0,1]$ executed in the framework of the $s$-th problem, consists of the following sequence of operations.

@@

	\item Compute the values $g_1(y^k),...,g_\nu(y^k)$, where the index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k)\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The trial terminates at the point as soon as the first violation of any constraint occurs. In the case when $y^k$ is a feasible one, the trial includes the computation of the values of all problem functions. In this situation, the index is set to $\nu=m+1$. The triplet
\[
y^s(x^k), \; \nu=\nu(x^k), \; z^k=g_\nu(y^s(x^k))
\]
is the outcome of the trial at the point $x^k$.

\item Determine the preimages $x^{k_s} \in [0,1]$, $1 \leq s \leq S$, of the point $y^k$ and interpret the trial executed at the point $y^k \in D$ as the execution of the trials at the $S$ points  $x^{k_1},...,x^{k_S}$ with the same outcomes
\[
 \nu(x^{k_1}) = ... = \nu(x^{k_S}) = \nu(x^k),
\]
\[
 g_\nu(y^1(x^{k_1})) = ... = g_\nu(y^S(x^{k_S})) = z^k.
\]

\item Inform the rest of processors about the trial outcome at the point $y^k$.

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. 
\Russian
Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области. При этом можно задавать объем допустимой области и  число ограничений, активных в точке минимума, тем самым усложняя или упрощая постановку задачи.

@@

The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta = 0.03\left\|b-a\right\|$. For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).

Computational experiments were carried out on Lobachevsky supercomputer. The node of supercomputer included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The CPU had 8 cores, i.e. each node had a total of 16 cores. The average time and the number of iterations, which were required to solve the problems of the series at various parallelization parameters are given in Tables 1 and 2. Here, \textit{Node(core)} are the numbers of employed  nodes and cores per node, respectively.

@@

The results demonstrate the speedup achieved when using the shared memory at a node as well as the distributed memory. In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.

@@

\Russian
В заключение отметим, что подавляющее большинство оптимизационных алгоритмов (за исключением методов, основанных на идеях случайного поиска) реализуют стратегию целенаправленного выбора. Указанная стратегия состоит в том, чтобы на основе анализа небольшой части вариантов исключить из дальнейшего рассмотрения многие заведомо не перспективные случаи и сосредоточить дальнейший поиск в подмножестве, содержащем лучший вариант.

Сложность моделируемых процессов и объектов может приводить к тому, что их характеристики не будут обладать свойством монотонности, что обуславливает недостаточность методов локального поиска для оценки наилучшего варианта. В таких задачах должны применяться методы глобального поиска, которые обеспечивают целенаправленность поиска за счет учета ограниченности изменения характеристик моделируемого объекта при ограниченных изменениях его параметров. Математическая формулировка этого факта может иметь форму Lipschitz condition, uniform H\"older condition и т.п. Именно к таким методам, способным решать сложные прикладные задачи оптимизации, относится рассмотренный в данной статье параллельный индексный алгоритм.

The parallel index method for solving constrained global optimization problems %considered in this paper 
offers the following possibilities:
\begin{itemize}
	\item to solve the initial problem directly, without the use of the penalty functions (thus eliminating the need to select the penalty coefficient and to solve a series of unconstrained problems with different penalty coefficients);
	\item to solve the problems where the values of the problem function are not defined everywhere (for example, the objective function values are undefined outside the feasible domain);
	\item to use the two-level parallelization scheme with the shared and distributed memory proposed earlier for the unconstrained optimization methods.
\end{itemize}

The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed. This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer.

\Russian
Перспективным направлением дальнейшей работы будет являться применение множественных отображений для редукции размерности в задачах многокритериальной оптимизации. В подобных задачах построение множества Парето требует решения целой серии скалярных задач. При их параллельном решении можно ожидать более высокого ускорения, чем при решении одной скалярной задачи.

@@

%\bibitem{Balandin2017}Balandin, D.V., Kogan, M.M.: Pareto-optimal generalized $H_2$-control and vibroprotection problems. Automation and Remote Control. 78(8), 1417--1429 (2017)
